Concurrent Programming with CXL

### 2020-06-15 book/paper.tex
Latest
Some text for barrier synchronization

\documentclass{report}
\usepackage{fullpage}
\usepackage[nottoc,numbib]{tocbibind}
\usepackage{graphicx}
\usepackage{framed}
\usepackage{tcolorbox}
\usepackage{verbatim}
\usepackage{fancyvrb}
\usepackage{comment}
\usepackage{amsmath}

@@

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/barrier-crop.pdf}
\end{center}
\caption{High-level state diagram specification of barrier synchronization
with three processes and three rounds.  The three numbers specify the round
number of each process.}
\label{fig:barrierdiagram}
\end{figure}

>>>>

\documentclass{report}
\usepackage{fullpage}
\usepackage[nottoc,notlot,notlof]{tocbibind}
\usepackage{graphicx}
\usepackage{framed}
\usepackage{tcolorbox}
\usepackage{verbatim}
\usepackage{fancyvrb}
\usepackage{comment}
\usepackage{amsmath}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/barrier-crop.pdf}
\end{center}
\caption{High-level state diagram specification of barrier synchronization
with three processes and two rounds.  The three numbers specify how many
rounds each process has completed.}
\label{fig:barrierdiagram}
\end{figure}

@@

Barrier synchronization is a problem that comes up in high-performance
parallel computing, used, among others, for scalable simulation.
Barrier synchronization is almost the opposite of critical sections:
the intention is to get a group of processes to run some code at the
same time.
More precisely, the processes run in rounds.  At the end of each round
there is a so-called \emph{barrier} where processes wait until all
processes have completed the round.
For example, in an iterative matrix algorithm, the matrix may be
cut up into fragments.  During a round, one process runs for each
fragment concurrently.  The next round is not allowed to start
until all processes have completed operations on their fragment.

Figure~\ref{fig:barrierdiagram} shows a high-level depiction with three
processes.  Initially all processes are in round 0.  Then each process
can start and finish the round.  However, none of the processes can
progress to the next round until all processes have finished the
current round.

Counting semaphores work well for implementing barrier synchronization.
Figure~\ref{fig:barrier} shows an example.  There is a 0-initialized
semaphore for each of the \texttt{N} processes.
Before process $i$ enters a round, it first sends a signal to every
other process and then waits until it receives a signal from
every other process.  Process $i$ sends a signal to process $j$
by incrementing \texttt{sema[$j$]} (using \texttt{V}),
while process $i$ waits for a signal be decrementing \texttt{sema[$i$]}
(using \texttt{P}).

The \texttt{round} array is kept to check the correctness of this
approach.  Each process increments its entry every time it enters
a round.  If the algorithm is correct, it can never be that two processes
are more than one round apart.
